### 2021-06-11 data-raw/Download.Rmd
Updated instructions

---
title: "Untitled"
author: "Dominique Kelly"
date: "6/3/2021"
output: html_document
---

```{r setup, include=FALSE}
#update.packages(ask = FALSE, checkBuilt = TRUE)
```

Check out this as an example: https://github.com/rudeboybert/fivethirtyeight/blob/master/data-raw/process_data_sets.R

Note:
* You'll be reading in an excel file, not a .csv
* Then make sure usethis::use_data() works
* Then do data cleaning

```{r}
usethis::use_data(MacLeish_Data_Plot_MC_01, overwrite = TRUE)
usethis::use_data(MacLeish_Data_Plot_MC_02, overwrite = TRUE)
```

```{r}
load_all()
```

```{r}
library(devtools)
library(tidyverse)
library(fs)
```

```{r}
plot_data <- as_tibble(MacLeish_Data_Plot_MC_01)
plot_data
```