### 2021-07-16 tests/testgraphs.Rmd
trivial change

---
title: "Tree Graphs Test"
author: "Dominique Kelly"
date: "7/9/2021"
output: html_document
---

```{r setup, include=FALSE}
library(macleish)
library(tidyverse)
library(ggplot2)
library(hrbrthemes)
library(viridis)
library(forcats)
```

```{r}
head(tree_diameter1)
```

```{r}
p <- tree_diameter1 %>%
  ggplot( aes(x=diameter)) +
    geom_histogram( binwidth=1, fill="#69b3a2", color="#e9ecef") +
    scale_x_continuous(limits = c(0,60)) +
    scale_y_continuous(limits = c(0,100))
p
```

```{r}
q <- tree_diameter1 %>%
  ggplot( aes(x=diameter, color=species, fill=species)) +
  geom_histogram(binwidth=50, color="#e9ecef", alpha=0.6, position = 'identity') +
  scale_fill_viridis(discrete=TRUE) +
  scale_color_viridis(discrete=TRUE) +
    theme(
      legend.position="none",
      panel.spacing = unit(0.1, "lines"),
      strip.text.x = element_text(size = 8)
    ) +
  #scale_x_continuous(limits = c(0,60)) +
  facet_grid(species~year, scales="free_y")
q
```

```{r}
x <- ggplot(tree_diameter1, aes(x = species)) +
  geom_bar(stat = "count") +
  theme(axis.text.x = element_text(angle = 60, hjust =1))
x
```

```{r}
z <- tree_diameter1 %>%
  mutate(position = fct_relevel(position,
                                "emergent", "canopy", "subcanopy", "sapling")) %>%
  ggplot(aes(x=position, y=diameter, fill=species)) +
  geom_bar(stat="identity", position="dodge") +
  scale_y_continuous(limits = c(0, 60))
z
#order of position - canopy, subcanopy, emergent, sapling 
#consider common names for species 
#add line graph for review
```

```{r}
w <- tree_diameter1 %>%
  ggplot( aes(x=diameter, y=species, fill=position)) +
  geom_boxplot() +
  scale_fill_viridis(discrete = TRUE, alpha=0.6, option="A") +
  scale_x_continuous(limits=c(0,60)) +
  facet_wrap(~year) +
  labs(x = "Diameter (cm)", y = "Species", fill = "Position", title = "Distribution of Diameter based on Different Species at Different Positions Between 2010-2015")
w
```

```{r}
w2 <- tree_diameter1 %>%
  ggplot(aes(x=diameter, y=species, fill=position)) +
  geom_boxplot() +
  scale_fill_viridis(discrete = TRUE, alpha=0.6, option="A") +
  scale_y_discrete(limits=c("Eastern Hemlock")) +
  coord_flip() +
  labs(x = "Diameter (cm)", y = "Species", fill = "Position", title = "Distribution of Diameter for Eastern Hemlock at Different Positions Between 2010-2015")
w2
```

```{r}
r <- tree_diameter2 %>%
  ggplot( aes(x=diameter, y=species)) +
  geom_boxplot() +
  geom_jitter(color="black", size=0.4, alpha=0.9) +
  scale_fill_viridis(discrete = TRUE, alpha=0.6, option="A") +
  scale_x_continuous(limits=c(0,60)) +
  facet_wrap(~year) +
  labs(x = "Diameter (cm)", y = "Species", title = "Distribution of Diameter based on Different Species Between 2009-2016")
r
```

```{r}
t <- tree_diameter1 %>%
  ggplot(aes(x=year, y=diameter, color=species)) +
  geom_line() + 
  geom_point()
t
```

# Any way to graph to focus on a specific year or plant species without butchering the data?